---
title: "checkpoint 2"
format: html
---

## Measles Dataset

```{r}
library("tidyverse")
cases_month <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')
cases_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
```

## Using existing columns
### Do the number of measles cases change depending on the time of year?
```{r}
monthlyStats <- function(monthly_measles) {
  tibble(mean_measles = mean(monthly_measles, na.rm = TRUE),
         median_measles = median(monthly_measles, na.rm = TRUE),
         max_measles = max(monthly_measles, na.rm = TRUE))
}

measles_by_month <- tibble()

for (i in 1:12) {
  monthly_measles <- cases_month |>
    filter(month == i) |> 
    pull(measles_total)
  
  measles_by_month <- bind_rows(measles_by_month, 
                                monthlyStats(monthly_measles) |>
                                  mutate(month = i))
}
measles_by_month <- measles_by_month |>
  select(month, mean_measles, median_measles, max_measles)
```
According to our resulting summaries, it appears the number of measles cases tends to
be the most during the beginning of the year (January - March), while the number of cases seems to be the lowest during the transition between summer and fall (August - October). However, it's important to keep in mind that these are summaries over the last 9 years, and include cases all over the world (and seasons/environment are different depending on location).
