---
title: "checkpoint 3"
author: Amanda Dong
format: html
---

## Country GDP API
Setting my API ninja key!! (code hidden)

```{r}
#| echo: false
my_key <- "GpzR11HILKq5OrK9rXSm7b9J0Fd4qvwfG6zu551Z"
```

load in packages and dataset:
```{r}
#| label: packages-setup
#| message: false
#| warning: false
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(stringr)
cases_month <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')
cases_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
```

**Example with Canada 2026**

To start, I am going to use the API to access the GDP data for Canada in 2026.
```{r}
base_url <- "https://api.api-ninjas.com/v1/gdp"
country <- "Canada"
year <- "2026"
url <- str_c(base_url, "?country=", country, "&year=", year)
canada_raw <- GET(url, add_headers("X-Api-Key" = my_key))
```

We now have the raw data, so I will change it into JSON format and parse in to a tidy dataframe.
```{r}
canada_data <- rawToChar(canada_raw$content)
canada_tidy <- fromJSON(canada_data)
canada_final <- canada_tidy %>% select(country, year, gdp_per_capita_nominal)
```


**Writing function**

Next, using the methods we used for Canada as a blueprint, I'll create a function to get the data for all countries.
```{r}
get_gdp_data <- function(country, year, key = my_key) {
  if (is.na(country) || is.na(year)) {
    return(NULL)
  }

  country <- str_replace_all(country, " ", "%20")
  base_url <- "https://api.api-ninjas.com/v1/gdp"
  url <- str_c(base_url, "?country=", country, "&year=", year)
  raw_data <- GET(url, add_headers("X-Api-Key" = my_key))
  json_data <- rawToChar(raw_data$content)
  tidy_data <- fromJSON(json_data)
  if (!is.data.frame(tidy_data) || nrow(tidy_data) == 0) {
    return(NULL)
  }
  if (!"gdp_per_capita_nominal" %in% names(tidy_data) || 
      is.na(tidy_data$gdp_per_capita_nominal)) {
    return(tibble(
      country = country,
      year = year,
      gdp_per_capita_nominal = NA
    ))
  }
  gdps <- tidy_data %>% select(country, year, gdp_per_capita_nominal)
}

```

Let's test the function using USA in 2019.
```{r}
usa <- get_gdp_data("United States", "2019")
get_gdp_data("hello", "yes")
```

Now we need to apply the function to all countries and years in our measles data.
```{r}
#| eval: false
apply_data <- cases_year %>% select(country, year) %>%
  filter(!is.na(country)) %>% filter(!is.na(year))

# results1 <- map2_dfr(.x = apply_data$country[1:500], .y = apply_data$year[1:500], .f = get_gdp_data)
# results2 <- map2_dfr(.x = apply_data$country[501:1000], .y = apply_data$year[501:1000], .f = get_gdp_data)
# results3 <- map2_dfr(.x = apply_data$country[1001:1500], .y = apply_data$year[1001:1500], .f = get_gdp_data)
# results4 <- map2_dfr(.x = apply_data$country[1501:2000], .y = apply_data$year[1501:2000], .f = get_gdp_data)
# results5 <- map2_dfr(.x = apply_data$country[2001:2380], .y = apply_data$year[2001:2380], .f = get_gdp_data)

results1$country_full <- countrycode(results1$country, origin = "iso3c", destination = "country.name")
results2$country_full <- countrycode(results2$country, origin = "iso3c", destination = "country.name")
results3$country_full <- countrycode(results3$country, origin = "iso3c", destination = "country.name")
results3[98, 4] <- "Pakistan"
results4$country_full <- countrycode(results4$country, origin = "iso3c", destination = "country.name")
results4[101:110, 4] <- "Montenegro"
results5$country_full <- countrycode(results5$country, origin = "iso3c", destination = "country.name")

GDP_APIdata <- bind_rows(results1, results2, results3, results4, results5)
write.csv(GDP_APIdata, "GDPdata.csv", row.names = FALSE)
```

**Brief description of new dataset**

We now have a new dataframe (which I've named "results") which contains 2380 rows with 3 variables. The 2380 rows correlate to the rows in "cases_year" (our original measles dataset) that have both country and year data. 2 of the variables in our new dataframe are country and year, while the last variable is "gdp_per_capita_nominal". As the name suggests, this column contains the nominal GDP per capita for that country, in that year. This was taken from API Ninja's dataset.